---
title: "week16_recap"
author: "Daniel Trielli"
date: "2023-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## Loading the packages

Run the codeblock below to load the packages we will need for this recap

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(tigris)
library(tidycensus)
library(corrr)
library(sf)
library(tidytext)
library(dplyr)
library(rvest)
```

## Load Data

[USAFacts compiled data on inspections on child care facilities throughout the country](https://usafacts.org/data-projects/childcare-inspections) - specifically, how many of those facilities are overdue for inspection. Let's take a look on the data they have on Maryland.

First, here the data we need to load:

```{r}
childcare_inspections <- read_csv('data/childcare_inspections.csv')
childcare_inspections$fips <- as.character(childcare_inspections$fips)
```

#### **Q1** The childcare data does not have the names of the counties. Join them with the md_counties data, that can be loaded in the block below. Then inspect the dataframe you created with this join and answer this: there are two columns with county names in the new dataframe. Which one is the best to be used? (hint: it has to do with homonyms)

```{r}
# This loads the county data
md_counties <- counties() |>  filter(STATEFP == "24")
```

```{r}
# Join the childcare data with the county data 
joined_data <- md_counties |> left_join(childcare_inspections, join_by(geoid == fips))
```

##### A1: We have to use the full name of the column "NAMELSAD" because it includes the full names for county and more importantly for Baltimore City. The other column writes baltimore twice for baltimore city and county, but we need to account of that potential duplicate if not careful. 

------------------------------------------------------------------------

#### **Q2** Create bar chart with the percentages of overdue childcare facility inspections per county to answer (You'll need to create a column for that calculation first). What is the county with the largest percentage of facilities with overdue inspections?

```{r}
# Calculate the percentage of overdue inspections per county
joined_data <- joined_data |>
  mutate(pct_overdue = ((overdue/total)*100))

# Create the chart with ggplot
joined_data |> 
  ggplot() +
  geom_bar(aes(x=reorder(namelsad, pct_overdue), weight=pct_overdue)) +
  coord_flip() +
  theme_minimal() +
  labs(
    title="Percent of overdue childcare inspections per Maryland county",
    x = "County",
    y = "Percent of overdue inspections",
) 
```

##### **A2: Montgomery County has the most overdue childcare facility inspections** 

------------------------------------------------------------------------

#### **Q3** Next, we want to see if the number of child care facilities - with overdue inspections or otherwise - is appropriate for each county. So we're going to calculate the rate of children per facility per county. To do that, you'll first need Census data to get the population of children under 5 for each county. Which county has the biggest rate of children per facility (i.e. they could use more facilities)?

```{r}
# First, grab data from the ACS. The ACS variable for children age under 5 is B06001_002 We want 2022 data.
under_5_kids <- get_acs(geography = "county",
              variables = c(medincome = "B06001_002"),
              state = "MD",
              year = 2022)
```

```{r}
# Rename the column name 'estimate', which has our population data, with something more useful, like 'children_pop'
under_5_kids <- under_5_kids |> 
  rename(children_pop = estimate)
```

```{r}
# Join the dataframes
acs_joined <- joined_data |> left_join(under_5_kids, join_by(geoid == GEOID))
# Calculate the rates for children per facility
acs_joined <- acs_joined |> 
  rename(county = namelsad) |>
  mutate(children_per_facility = children_pop/total)

acs_joined |>
  select(county, children_per_facility) |>
  arrange(desc(children_per_facility))

acs_joined
```

##### **A3: Garrett County has the most children per facility.** 

------------------------------------------------------------------------

#### **Q4** Make a map of the results for question 3. Are there any areas of the state that are more lacking in childcare facilities?

```{r}
ggplot() +
  geom_sf(data=acs_joined, aes(fill=children_per_facility)) +
  theme_minimal() 
```

##### **A4**: Some of the coutnies with the lowest number of children per facility are also some of the lesser populated, rural Eastern Maryland counties. The urban counties seem to have more children per facility but then the clear exception of Garrett County in the western edge of the state. 

------------------------------------------------------------------------

#### **Q5** Now let's explore if we can find patterns in under-served areas, particularly when it comes to median income of the population. First, join the childcare data median income data from the Census. Then, create a scatterplot and perform a cor.test(), and answer: is there is any correlation between income and rate of children per facilities? What makes you conclude that?

```{r}
# Again, grab data from the ACS. The ACS variable for median income is B19013_001. We want 2022 data again.
median_income <- get_acs(geography = "county",
              variables = c(medincome = "B19013_001"),
              state = "MD",
              year = 2022) |> clean_names()
```

```{r}
# Rename the column name 'estimate', which has our income data, with something more useful, like 'median_income'
median_income <- median_income |> 
  rename(median_income = estimate) 
```

```{r}
# Join the dataframes
income_joined <- acs_joined |> left_join(median_income, join_by(geoid))
```

```{r}
# Create a scatterplot with median income and rate of children per facility
income_joined |>
  ggplot() +
  geom_point(aes(x=median_income,y=children_per_facility)) +
  geom_smooth(aes(x=median_income,y=children_per_facility), method="lm")

```

```{r}
# Create a correlation test.
cor.test(income_joined$median_income, income_joined$children_per_facility)
```

##### A5: There is a slight negative correlation between the two variables, but it does not appear statistically significant given the p-value is greater than 0.05

------------------------------------------------------------------------

#### **Q6** Finally, let's do some text analysis. We another dataset, with inspection reports from informal child care providers in the state. This dataset has an informative column of initial findings: a column that describes issues found in these providers, if any. Your goal is to create bigrams for this column and find common issues that are mentioned in these findings. And then answer: what kind of issues listed there you think are interesting? They don't have the most frequent.

```{r}
# Read the reports data
reports <- read_csv('data/childcare-informal-provider-inspections.csv') |> clean_names()
```

```{r}
# Create a stopword dictionary (feel free to include more words)
data("stop_words")

stop_words <- stop_words |> 
  add_row(word = "H&S") |> 
  add_row(word = "with") |> 
  add_row(word = "local") |>
  add_row(word = "prepared") |>
  add_row(word = "maryland") |>
  add_row(word = 'u.s') |> 
  add_row(word = 'md') |> 
  add_row(word = 'court') |> 
  add_row(word = "original") |>
  add_row(word = "account") |>
  add_row(word = "respondent") |> 
  add_row(word = "client") 
```

```{r}
# Unnest tokens into bigrams, removing stop words and repasting bigrams, list bigrams
unique_words <- reports |> select(initial_findings) |>
  unnest_tokens(word, initial_findings)

unique_words <- unique_words |>
  anti_join(stop_words)

reports |>
  unnest_tokens(bigram, initial_findings, token = "ngrams", n = 2) |>
  separate(bigram, c("word1", "word2"), sep = " ") |>
  filter(!word1 %in% stop_words$word) |>
  filter(!word2 %in% stop_words$word) |>
  mutate(bigram = paste(word1, word2, sep=" ")) |>
  group_by(bigram) |>
  tally(sort=TRUE) |>
  top_n(10)
```

##### A6: There appear to be some significant repeated issues here with regard to missing an "emergency bag" albeit unclear what that means, emergency seems important. Also issues with smoke detectors is a huge problem. Same with outlet coverings because that's a fire hazard. 
