---
title: "Reverse Engineering Project"
author: "Matt Cohen, "
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story: [AP's story on missing kids at schools](https://projects.apnews.com/features/2023/missing-children/index.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data
population_by_age <- read_csv('sb152xr1685/population-by-age.csv') |> clean_names()
private_enrollment_by_grade <- read_csv('sb152xr1685/private-enrollment-by-grade.csv') |> clean_names()
public_enrollment_by_grade <- read_csv('sb152xr1685/public-enrollment-by-grade.csv') |> clean_names()
summary_file <- read_csv('sb152xr1685/enrollment-changes.csv') |> clean_names()


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: Overall, public school enrollment fell by over 700,000 students between the 2019-2020 and 2021-2022 school years in the 21 states plus Washington, D.C., that provided the necessary data.
-   **Analysis summary**: After several attempts to figure out how to sort through some confusingly structured data, we figured out how the AP reached the claim in 700,000 students gone from public school enrollment from 2019-2020 to 2021-22. After mutating and filtering the data, we produced a summary that showed total enrollment in the 21 states plus DC that the AP used for its analysis, based on the above sentence. We then subtracted the 2021-22 total enrollment from the 2019-20 and got 704,593. That lines up with the sentence in the AP style saying enrollment declined by "over 700,000"

```{r}
# Put code to reverse engineer sentence here

filtered_states <- public_enrollment_by_grade |> 
  filter(grade == "k_12_total") |>
  mutate(
    counted_state = case_when(
        str_detect(st, "CA") ~ "TRUE",
        str_detect(st, "NY") ~ "TRUE",
        str_detect(st, "LA") ~ "TRUE",
        str_detect(st, "NC") ~ "TRUE",
        str_detect(st, "WA") ~ "TRUE",
        str_detect(st, "CO") ~ "TRUE",
        str_detect(st, "GA") ~ "TRUE",
        str_detect(st, "MN") ~ "TRUE",
        str_detect(st, "WI") ~ "TRUE",
        str_detect(st, "NH") ~ "TRUE",
        str_detect(st, "MA") ~ "TRUE",
        str_detect(st, "MT") ~ "TRUE",
        str_detect(st, "DE") ~ "TRUE",
        str_detect(st, "ND") ~ "TRUE",
        str_detect(st, "DC") ~ "TRUE",
        str_detect(st, "SD") ~ "TRUE",
        str_detect(st, "SC") ~ "TRUE",
        str_detect(st, "NE") ~ "TRUE",
        str_detect(st, "RI") ~ "TRUE",
        str_detect(st, "TN") ~ "TRUE",
        str_detect(st, "PA") ~ "TRUE",
        str_detect(st, "FL") ~ "TRUE",
        .default = "FALSE",
    )
  )


pre_covid_publics <- filtered_states |> 
  filter(sch_year == "2019-2020") |>
  filter(counted_state == "TRUE") |>
  summarize(total = sum(enrollment, na.rm = TRUE))

post_covid_publics <- filtered_states |> 
  filter(sch_year == "2021-2022") |>
  filter(counted_state == "TRUE") |>
  summarize(total = sum(enrollment, na.rm = TRUE))

public_change <- pre_covid_publics - post_covid_publics 

public_change

# Display results of code below this codeblock

```

### Sentence 2

-   **Sentence text**: Those states saw private-school enrollment grow by over 100,000 students.
-   **Analysis summary**: This is a follow up on the previous sentence, but this one discusses private schools. The code to solve this one is pretty much identical, but did take checking some slightly altered column names compared to the public school data. When subtracting the 2021-22 enrollment from the 2019-20, we get -102,847. That in fact means that there was an increase of just over 100,000 in the total private school enrollment. That exactly lines up with the AP story.

```{r}
# Put code to reverse engineer sentence here
filter_privates <- private_enrollment_by_grade |> 
  filter(grade == "k_12_total") |>
  mutate(
    counted_state = case_when(
        str_detect(state, "CA") ~ "TRUE",
        str_detect(state, "NY") ~ "TRUE",
        str_detect(state, "LA") ~ "TRUE",
        str_detect(state, "NC") ~ "TRUE",
        str_detect(state, "WA") ~ "TRUE",
        str_detect(state, "CO") ~ "TRUE",
        str_detect(state, "GA") ~ "TRUE",
        str_detect(state, "MN") ~ "TRUE",
        str_detect(state, "WI") ~ "TRUE",
        str_detect(state, "NH") ~ "TRUE",
        str_detect(state, "MA") ~ "TRUE",
        str_detect(state, "MT") ~ "TRUE",
        str_detect(state, "DE") ~ "TRUE",
        str_detect(state, "ND") ~ "TRUE",
        str_detect(state, "DC") ~ "TRUE",
        str_detect(state, "SD") ~ "TRUE",
        str_detect(state, "SC") ~ "TRUE",
        str_detect(state, "NE") ~ "TRUE",
        str_detect(state, "RI") ~ "TRUE",
        str_detect(state, "TN") ~ "TRUE",
        str_detect(state, "PA") ~ "TRUE",
        str_detect(state, "FL") ~ "TRUE",
        .default = "FALSE",
    )
  )


pre_covid_privates <- filter_privates |> 
  filter(sch_year == "2019-2020") |>
  filter(counted_state == "TRUE") |>
  summarize(total = sum(enrollment, na.rm = TRUE))

post_covid_privates <- filter_privates |> 
  filter(sch_year == "2021-2022") |>
  filter(counted_state == "TRUE") |>
  summarize(total = sum(enrollment, na.rm = TRUE))

private_change <- pre_covid_privates - post_covid_privates

private_change



# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: The number of children registered for home-schooling surged by around 184,000.
-   **Analysis summary**: The missing student data came from a compilation of enrollment changes at public school, private school and home school. We previously calculated public and private schools, now we complete it with home schools students. We took the same approach that was used for previous questions, but this data set was incorporated into the summary file for the data instead of in its own table. So this took an extra step of filtering. Once that was completed, we were able to calculate the exact difference in enrollment by summarizing the sums of the 2021-22 school year enrollment and 2019-20 and then finding the difference. That math resulted in the same number the AP story had: around 184,000.

```{r}
# Put code to reverse engineer sentence here
homeschool <- summary_file |> filter(enrollment_type == "Homeschool enrollment") |> mutate(
    counted_state = case_when(
        str_detect(state_abbrev, "CA") ~ "TRUE",
        str_detect(state_abbrev, "NY") ~ "TRUE",
        str_detect(state_abbrev, "LA") ~ "TRUE",
        str_detect(state_abbrev, "NC") ~ "TRUE",
        str_detect(state_abbrev, "WA") ~ "TRUE",
        str_detect(state_abbrev, "CO") ~ "TRUE",
        str_detect(state_abbrev, "GA") ~ "TRUE",
        str_detect(state_abbrev, "MN") ~ "TRUE",
        str_detect(state_abbrev, "WI") ~ "TRUE",
        str_detect(state_abbrev, "NH") ~ "TRUE",
        str_detect(state_abbrev, "MA") ~ "TRUE",
        str_detect(state_abbrev, "MT") ~ "TRUE",
        str_detect(state_abbrev, "DE") ~ "TRUE",
        str_detect(state_abbrev, "ND") ~ "TRUE",
        str_detect(state_abbrev, "DC") ~ "TRUE",
        str_detect(state_abbrev, "SD") ~ "TRUE",
        str_detect(state_abbrev, "SC") ~ "TRUE",
        str_detect(state_abbrev, "NE") ~ "TRUE",
        str_detect(state_abbrev, "RI") ~ "TRUE",
        str_detect(state_abbrev, "TN") ~ "TRUE",
        str_detect(state_abbrev, "PA") ~ "TRUE",
        str_detect(state_abbrev, "FL") ~ "TRUE",
        .default = "FALSE",
    )
  )

home_change <- homeschool |>
  filter(counted_state == "TRUE") |>
  summarize(total = (sum(school_year_2021_22, na.rm=TRUE))-(sum(school_year_2019_20, na.rm=TRUE)))

home_change

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: But the data showed 230,000 students who were neither in private school nor registered for home-school. Their absences could not be explained by population loss, either – such as falling birth rates or families who moved out of state.
-   **Analysis summary**: We were able to get the same result as the AP through a several step process that included manipulating rows in order to make the data join-able and parsing out the 2019-20 data from the 2021-22 data in order to make it possible to add and subtract across the two school years. We then joined all the data and created a similar data set to what the AP used in the summary file to create a residual file (adding the public school change, private school change and home school change before subtracting that from the total population change in school age children). That total got us the same amount as the AP wrote, 234,213 missing students.

```{r}
# Put code to reverse engineer sentence here
school_population <- read_csv('sb152xr1685/school_population.csv') |> clean_names()

counted_population <- school_population |> mutate(
    counted_state = case_when(
        str_detect(state, "California") ~ "TRUE",
        str_detect(state, "New York") ~ "TRUE",
        str_detect(state, "Louisiana") ~ "TRUE",
        str_detect(state, "North Carolina") ~ "TRUE",
        str_detect(state, "Washington") ~ "TRUE",
        str_detect(state, "Colorado") ~ "TRUE",
        str_detect(state, "Georgia") ~ "TRUE",
        str_detect(state, "Minnesota") ~ "TRUE",
        str_detect(state, "Wisconsin") ~ "TRUE",
        str_detect(state, "New Hampshire") ~ "TRUE",
        str_detect(state, "Massachusetts") ~ "TRUE",
        str_detect(state, "Montana") ~ "TRUE",
        str_detect(state, "Delaware") ~ "TRUE",
        str_detect(state, "North Dakota") ~ "TRUE",
        str_detect(state, "District of Columbia") ~ "TRUE",
        str_detect(state, "South Dakota") ~ "TRUE",
        str_detect(state, "South Carolina") ~ "TRUE",
        str_detect(state, "Nebraska") ~ "TRUE",
        str_detect(state, "Rhode Island") ~ "TRUE",
        str_detect(state, "Tennessee") ~ "TRUE",
        str_detect(state, "Pennsylvania") ~ "TRUE",
        str_detect(state, "Florida") ~ "TRUE",
        .default = "FALSE",
    )
  ) 


edit_private <- filter_privates |> mutate(
  state = case_when(
        str_detect(state, "CA") ~ "CALIFORNIA",
        str_detect(state, "NY") ~ "NEW YORK",
        str_detect(state, "LA") ~ "LOUISIANA",
        str_detect(state, "NC") ~ "NORTH CAROLINA",
        str_detect(state, "WA") ~ "WASHINGTON",
        str_detect(state, "CO") ~ "COLORADO",
        str_detect(state, "GA") ~ "GEORGIA",
        str_detect(state, "MN") ~ "MINNESOTA",
        str_detect(state, "WI") ~ "WISCONSIN",
        str_detect(state, "NH") ~ "NEW HAMPSHIRE",
        str_detect(state, "MA") ~ "MASSACHUSETTS",
        str_detect(state, "MT") ~ "MONTANA",
        str_detect(state, "DE") ~ "DELAWARE",
        str_detect(state, "ND") ~ "NORTH DAKOTA",
        str_detect(state, "DC") ~ "DISTRICT OF COLUMBIA",
        str_detect(state, "SD") ~ "SOUTH DAKOTA",
        str_detect(state, "SC") ~ "SOUTH CAROLINA",
        str_detect(state, "NE") ~ "NEBRASKA",
        str_detect(state, "RI") ~ "RHODE ISLAND",
        str_detect(state, "TN") ~ "TENNESSEE",
        str_detect(state, "PA") ~ "PENNSYLVANIA",
        str_detect(state, "FL") ~ "FLORIDA",
        .default = "FALSE",
    )
  ) 

Y19_privates <- edit_private |> filter(counted_state == "TRUE", sch_year == "2019-2020")

Y21_privates <- edit_private |> filter(counted_state == "TRUE", sch_year == "2021-2022")

join_privates <- Y19_privates |> left_join(Y21_privates, join_by(state)) 

joined_privates <- join_privates |> mutate(private_change = enrollment.y - enrollment.x)

Y19_publics <- filtered_states |> filter(counted_state == "TRUE", sch_year == "2019-2020")

Y21_publics <- filtered_states |> filter(counted_state == "TRUE", sch_year == "2021-2022")

join_publics <- Y19_publics |> left_join(Y21_publics, join_by(state))

joined_publics <- join_publics |> mutate(public_change = enrollment.y - enrollment.x)

caps_homeschool <- homeschool |> mutate(state = str_to_upper(state))

counted_population_caps <- counted_population |> mutate(state = str_to_upper(state))

all_joined_data <- joined_privates |> 
  left_join(joined_publics, join_by(state)) |> 
  left_join(caps_homeschool, join_by(state)) |>
  left_join(counted_population_caps, join_by(state))

all_joined_data <- all_joined_data |> select(private_change, public_change, total_change, change_in_school_aged_population_5_17)

missing <- all_joined_data |> mutate(residual = (private_change + public_change + total_change - change_in_school_aged_population_5_17)) |> arrange(residual)

missing |> summarize(sum(residual))


# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: The following sentence is NOT in the story: If we filtered out the counted states, what would be the total number of students missing across all available data in this notebook?
-   **Analysis summary**: The answer is about 290,000 missing students, or about 60,000 more than when we only used the counted states. A key for this data is to note this still is not all 50 states plus DC. Only the public school data and population data includes information for all 50 states plus DC. The private school data and home school data does not. So the final answer we have is not without holes. But it is the sum of the residuals for all of the states in the data set, albeit not 50 plus DC. The main tactic to reach this number, similar to sentence four, was to separate out the 2019-20 data and 2021-22 data in the public and private schoold data so we could subtract and the get the difference. From there, we joined all the data together and got the residual. The key difference here is that we just removed the filter we had previously been using.

```{r}
# Put code to reverse engineer sentence here
public_enrollment_by_grade_19 <- public_enrollment_by_grade |> filter(sch_year == "2019-2020", grade == "k_12_total")
public_enrollment_by_grade_21 <- public_enrollment_by_grade |> filter(sch_year == "2021-2022", grade == "k_12_total")
private_enrollment_by_grade_19 <- private_enrollment_by_grade |> filter(sch_year == "2019-2020", grade == "k_12_total")
private_enrollment_by_grade_21 <- private_enrollment_by_grade |> filter(sch_year == "2021-2022", grade == "k_12_total")
all_50_homeschool <- summary_file |> filter(enrollment_type == "Homeschool enrollment") 

public_enrollment <- public_enrollment_by_grade_19 |> left_join(public_enrollment_by_grade_21, join_by(state))
private_enrollment <- private_enrollment_by_grade_19 |> left_join(private_enrollment_by_grade_21, join_by(state))

public_enrollment <- public_enrollment |> mutate(public_change = enrollment.y - enrollment.x)
private_enrollment <- private_enrollment |> mutate(private_change = enrollment.y - enrollment.x)

private_enrollment <- private_enrollment |> mutate(state = case_when(
        str_detect(state, "CA") ~ "CALIFORNIA",
        str_detect(state, "CO") ~ "COLORADO",
        str_detect(state, "NY") ~ "NEW YORK",
        str_detect(state, "HI") ~ "HAWAII",
        str_detect(state, "IN") ~ "INDIANA",
        str_detect(state, "IA") ~ "IOWA",
        str_detect(state, "KS") ~ "KANSAS",
        str_detect(state, "MD") ~ "MARYLAND",
        str_detect(state, "MI") ~ "MICHIGAN",
        str_detect(state, "LA") ~ "LOUISIANA",
        str_detect(state, "NC") ~ "NORTH CAROLINA",
        str_detect(state, "WA") ~ "WASHINGTON",
        str_detect(state, "CO") ~ "COLORADO",
        str_detect(state, "GA") ~ "GEORGIA",
        str_detect(state, "MN") ~ "MINNESOTA",
        str_detect(state, "WI") ~ "WISCONSIN",
        str_detect(state, "NH") ~ "NEW HAMPSHIRE",
        str_detect(state, "MA") ~ "MASSACHUSETTS",
        str_detect(state, "MT") ~ "MONTANA",
        str_detect(state, "DE") ~ "DELAWARE",
        str_detect(state, "ND") ~ "NORTH DAKOTA",
        str_detect(state, "DC") ~ "DISTRICT OF COLUMBIA",
        str_detect(state, "SD") ~ "SOUTH DAKOTA",
        str_detect(state, "SC") ~ "SOUTH CAROLINA",
        str_detect(state, "NE") ~ "NEBRASKA",
        str_detect(state, "RI") ~ "RHODE ISLAND",
        str_detect(state, "TN") ~ "TENNESSEE",
        str_detect(state, "PA") ~ "PENNSYLVANIA",
        str_detect(state, "FL") ~ "FLORIDA",
        str_detect(state, "OH") ~ "OHIO",
        str_detect(state, "TX") ~ "TEXAS",
        str_detect(state, "NJ") ~ "NEW JERSEY",
        str_detect(state, "WV") ~ "WEST VIRGINIA",
        str_detect(state, "CT") ~ "CONNECTICUT",
        str_detect(state, "NV") ~ "NEVADA",
  )
)

all_50_joined_data <- public_enrollment |> 
  left_join(private_enrollment, join_by(state)) |> 
  left_join(caps_homeschool, join_by(state)) |>
  left_join(counted_population_caps, join_by(state))

all_50_missing <- all_50_joined_data |> mutate(residual = (private_change + public_change + total_change - change_in_school_aged_population_5_17)) |> arrange(residual)

all_50_missing |> summarize(missing = sum(residual, na.rm=TRUE))




# Display results of code below this codeblock

```

-30-
