---
title: "lab_05"
author: "derek willis"
date: "2023-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   Tabula

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
library(janitor)
library(lubridate)
```

## Get Our PDF

We'll be working with the [911 overdose calls from Baltimore County](https://drive.google.com/file/d/1qkYuojGF_6WKFr5aNQxmewDzcKyOiJFr/view?usp=share_link). You'll want to download it to a place you'll remember (like your Downloads folder, or the labs folder in your repository). The goal is to extract the tables within it, export that to a CSV file, load it into RStudio and ask some questions.

## Extract Data from PDF Using Tabula

Start Tabula, then go to <http://127.0.0.1:8080/> in your browser. Click the "Browse" button and find the PDF file and click "open", and then click the "Import button" in Tabula. This will take a few seconds or longer.

This PDF has a single table spread over multiple pages to extract. We're going to make a single dataframe from this table, exporting it to a CSV file that you will load into R. In Tabula, highlight the table and click the "Preview & Export Extracted Data" button. You may want to play with including or excluding the column headers - YOU SHOULD HAVE FIVE COLUMNS OF DATA.

Save the CSV (it should be called `tabula-Baltimore County; Carey, Samantha log OD.csv` by default) to your lab_05/data folder.

From there, you will need to read in the data, and add or fix headers if necessary. You can choose to include the headers from the PDF in your exported CSV files OR to exclude them and add them when importing. `read_csv` allows us to do this ([and more](https://readr.tidyverse.org/reference/read_delim.html)).

## Load and clean up the data in R

You will need to read in and clean up the data so that it can be used for analysis. By "clean" I mean the column headers should not contain spaces and they should have meaningful names, not "x1" or something similar. How you do that is up to you, but you can use select() with or without the minus sign to include or exclude certain columns. You also can use the `rename` function to, well, rename columns. Importantly, you'll need to ensure that any columns containing a date actually have a date datatype. Our friend `lubridate` can help with this.

```{r}
calls <- 
  read_csv("lab_five_data.csv", col_names=FALSE) |> 
  clean_names() |> 
  rename(date = x1, time = x2, case_number = x3, evtyp = x4, location = x5) |> 
  mutate(date = mdy(date))

calls
```

## Answer questions

Q1. Write code to generate the number of calls that occurred on each date. Which date in 2022 had the most overdose calls, and how many? Look at the total number of rows in your result and explore the range of dates - based on your result, do you believe there are any days with no overdose calls at all? Explain why or why not.

**A1. Two dates in 2022 had 23 total calls. They were July 14 and October 4. There appears to have been a call every day plus an extra day because there are more than 365 rows of dates.**

```{r}
calls |>
  group_by(date) |>
  summarize(total = n()) |>
  arrange(desc(total))
```

Q2. You want to understand if there's a pattern in the day of the week that overdose calls are made. Add a column to your dataframe that displays what day of the week each date represents. You should search for how to do that using lubridate. Then write code to calculate the number of calls for each day of the week, and add a column to that result that calculates the percentage of all calls that occurred on each day of the week (so you want a dataframe with the day of the week, total number of calls and the percentage of calls on that day out of the total number of all calls). Describe your findings to me.

**A2. The most calls, probably without much surprise, were made over the weekend between Friday-Sunday. Of the weekly calls, 15.5% come on Saturday, 15.1% come on Sundays and 15% come on Fridays.**

```{r}
days <- calls |> mutate(day = wday(date, label=TRUE))

days |> 
  group_by(day) |>
  summarize(total_calls = n()) |>
  arrange(desc(total_calls)) |>
  mutate(percent_calls = (total_calls/4112)*100)
 

```

Q3. Now let's look at locations. Which ones have the most calls? How would you describe them (feel free to search for more information on them)? Is there anything about the structure of the original data that might make you less confident in the counts by location or date?

**A3. The most frequent location is 4540 Silver Spring Road which appears to be in Perry Hall. That is Baltimore County. My concern though would be the next two highest frequency locations are both police precints for Baltimore County PD. So that doesn't really give information on where exactly the 911 calls are taking place but more so a neighborhood where people end up in the police precinct. The biggest trend that is notable in this data is every single one of the top 10 most frequent locations for calls comes from Baltimore County. Is that shocking? Not really. But that level of frequency confirms an already believed to exist problem.**

```{r}
calls |> 
  group_by(location) |>
  summarize(total_calls = n()) |>
  arrange(desc(total_calls))
```

Q4. What's the best story idea or question you've seen as a result of the work you've done in this lab?

**A4. Given that Baltimore County has clearly proven to have most frequent 911 calls, I'd like to see if we could organize this data by county. For the sake of this, let's assume the data already included the county of each location so that it was easier to filter quickly for instead of having to look them all up. If I had that info, I would compare how many more calls came from Baltimore County than other places.**
