---
title: "lab_01"
author: "Derek Willis"
adapted by: "Daniel Trielli"
date: "2024-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your class reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the UMD courses data.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
umd_courses <- read_rds("lab_01.rds")
```

## Answer questions

### **Q1** Which department has the highest mean (average) number of seats for its classes? You need to show the department with the highest average first. Which one is the most surprising/interesting to you?

```{r}
umd_courses |>
  group_by(department) |>
  summarise(
    mean_seats = mean(seats)
  ) |>
  arrange(desc(mean_seats))

```

**A1 The "engineering, cooperative education" department has the highest mean number of seats per class.**

------------------------------------------------------------------------

### **Q2** Find out the greatest number of times any individual class has been offered in our dataset. You must show the greatest number of classes first.

```{r}
umd_courses |>
  group_by(title) |>
  summarise(
    count_classes=n()
  ) |>
  arrange(desc(count_classes))
```

**A2 Doctrinal Dissertation Research has been offered 1,303 times over the course of this data set, making it the most frequently offered class at UMD.**

------------------------------------------------------------------------

### **Q3** Count the number of courses for each instructor and department. Which instructor has taught the most?

```{r}
umd_courses |>
  group_by(department)|>
  summarise(
    count_classes=n()
  ) |>
  arrange(desc(count_classes))


umd_courses |>
  group_by(instructors)|>
  summarise(
    count_classes=n()
  ) |>
  arrange(desc(count_classes))



```

**A3 Elsa Barkley Brown teaches 73 classes, which is the most for any one professor. The Business and Management department is the dept. with the most course offerings.**

### **Q4** Given the results for Q3, what would have made it easier to find the answer? You don't have to know *how* to accomplish this; I want you to describe in words what would make the answer easier to see.

**A4 I want to be able to combine it into one table so that I could see what dept. each instructor is in. It also would be easier to filter out the number of classes that have an "N/A" or "TBA" listing for instructor.**

------------------------------------------------------------------------

### **Q5** Take one of the questions you listed in the final question of the pre_lab and try to answer it here in code. If you don't know how to fully answer it, write what code you can and describe in words what else you would need to do. If you got an answer, is it interesting?

```{r}
umd_courses |>
  group_by(instructors)|>
    summarize(
      avg_seats = mean(seats)
    ) |>
  arrange(desc(avg_seats))

```

**A5 This data alone doesn't answer my pre-lab questions, I asked about comparing it to enrollment data and figuring out how many of the available seats are actually filled. If we had that data, we could us code to calcuate the ratio of students to instructors, or subtraction to figure out the number of unused seats. Because I can't calculate that with what we have, I tried to instead find the ratio of instructors to seats. The code I wrote somewhat works, but in doing so learned how many classes have a giant number of seats because there are large groups of profs grouped together. Nor is the mean number of seats by professor the best way to calculate the ratio of seats to professor, but it's the closest apporximation I know how to do. I don't know how to untangle all the large groups of instructors leading to some averages approaching 3,000 seats.**
