---
title: "week7_recap"
author: "Daniel Trielli"
date: "2023-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## Loading the packages

Run the codeblock below to load the packages we will need for this recap

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load Data

Run the codeblock below to load the data.

```{r}
earthquakes <- read_csv('https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/all_month.csv')

#Setting time column as datetime
#earthquakes <- earthquakes |> mutate(time = as_datetime(time))
```

#### Answer the questions below

Most questions have a code block and a space for an answer below. Write the code you think is necessary and, in the answer space, write out what you did and what was the result.

------------------------------------------------------------------------

#### **Q1** Do you notice anything different with this read_csv()? Particularly in the file name?

**A1: It's a URL!**

------------------------------------------------------------------------

#### **Q2** How many records there are there in this dataset? What do they mean and what useful information we can gather from it, looking at the columns?

**A2: Magnitude, where the earthquake is, what time it occurred**

------------------------------------------------------------------------

#### **Q3** How do I reorganize this list to see the ones that are the deepest first? What is the depth that shows up for the deepest one, and its magnitude?

```{r}
earthquakes |>
  arrange(desc(depth))

```

**A3: 628.362 KM below sea level. It was a 4.6 magnitude earthquake**

------------------------------------------------------------------------

#### **Q4** I just want to see the earthquakes with a magnitude larger than 6. How do I do that? And how many are there that fit this criteria?

```{r}
earthquakes |> filter(mag > 6)
```

**A4: There are six earthquakes in the last hour with magnitude greater than 6.**

------------------------------------------------------------------------

#### **Q5** What about if I want to see earthquakes that have both a magnitude larger than 6 and a depth smaller than 20? How many are there in the data set that fit [both]{.underline} these criteria?

```{r}
earthquakes |> filter(mag>6 & depth<20)
```

**A5: There are four examples that fit this example.**

------------------------------------------------------------------------

#### **Q6** What about if I want to see earthquakes that either have a magnitude larger than 6 and a depth smaller than 20? How many are there in the data set that fit [either]{.underline} these criteria?

```{r}
earthquakes |> filter(mag>6 | depth<20)
```

**A6: There are 7,475 examples that cover either option.**

------------------------------------------------------------------------

#### **Q7** I'm interested in finding earthquakes that took place in Alaska. Which column should I look at? How do I use it to find all the earthquakes in Alaska? How many earthquakes took place there?

```{r}
earthquakes |> filter(str_detect(place, "Alaska"))
```

**A7: There are 2,168 earthquakes detected in Alaska counted in the last hour.**

------------------------------------------------------------------------

#### **Q8** I notice that there is a column called 'type', that seems to have different kinds of tremors. What is the best way to find what are all the possible types of tremors, and counting how many of each there are in this data set? What are the first two most common types of tremors in this data set?

```{r}
earthquakes |> group_by(type) |> summarize(total = n())
```

**A8: Earthquakes and quarry blasts are the most common types in the data set.**

------------------------------------------------------------------------

#### **Q9** How do I calculate the average depth of each type of tremor? What is the average depth of the earthquake type in this data set? Is there anything that seems unusual or surprising in this finding?

```{r}
earthquakes |> group_by(type) |> summarize(avg = mean(depth)) |> arrange(desc(avg))
```

**A9: The average earthquake depth is 19.4 KM below sea level. What is interesting in this data set is that there are explosions and quarry blasts taking place above sea level but being picked up as a tremor.**

------------------------------------------------------------------------

#### **Q10** I'm interested, in the future, to see the hours in which earthquakes happen. How can I extract the hour from the time column?

```{r}

```

**A10:**

------------------------------------------------------------------------

#### **Q11** I want to make a note of all the records in this data set that I consider serious. For my purposes, I'm thinking that all tremors that have a magnitude that is larger than 3 are serious. How do I automatically create a new column making that distinction?

```{r}
serious_quakes <- earthquakes |>
  mutate (serious = case_when(
    mag > 3 ~ "S",
    !mag > 3 ~ "N/A"
  )
)

serious_quakes
```

**A11:**

------------------------------------------------------------------------

#### **Q12** I have no idea how earthquakes work and I'm interested in seeing if there is a particular time of day in which serious earthquakes happen. How can I see that condensed in a table with all the hours in a day and all the serious earthquakes in each hour? What is the hour with fewer serious earthquakes and the one with the most serious earthquakes?

```{r}

```

**A12**:

------------------------------------------------------------------------

#### EXTRA Q What other questions are you interested in and how would you ask them in R?

```{r}
earthquakes |>
  group_by(locationSource) |>
  summarize(total_tremors = n()) |>
  arrange(desc(total_tremors))
  
```

```{r}
earthquakes |>
 filter(type == "explosion") |> group_by(locationSource) |> summarize(total_blasts = n()) |> arrange(desc(total_blasts))
```

```{r}
earthquakes |>
 filter(type == "earthquake") |> group_by(locationSource) |> summarize(total_quakes = n()) |> arrange(desc(total_quakes))
```
