---
title: "lab_04_ai"
author: "Daniel Trielli"
date: "2024-09-26"
output: html_document
---

## Setup

Let's load the necessary libraries and, using your API key, setup your credentials:

```{r}
library(axolotr)
create_credentials(GROQ_API_KEY = "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")
```

See that "Please restart your R session for changes to take effect."? Go ahead and do that; you'll need to rerun the `library()` function for axolotr, and let's load tidyverse while we're at it.\

```{r}
library(axolotr)
library(tidyverse)
```

## Testing

Let's make sure that worked. We'll be using the [Llama 3.1 model released by Meta](https://ai.meta.com/blog/meta-llama-3-1/).

```{r}
groq_response <- axolotr::ask(
  prompt = "Give me a 100-word pitch for a new Lilo and Stitch sequel",
  model = "llama-3.1-8b-instant"
)

groq_response
```

Did that work?

Yes!

## Q1. Turning unstructured information into data: let's take [this article](https://www.nytimes.com/2024/09/27/arts/maggie-smith-dead.html) about the death of Dame Maggie Smith and try to extract information from it. Your goal is to have Groq create a table with the films and TV shows mentioned in this news article, and extract the following information about them: title, year, role, director, co-stars, and awards. From the table that it created, answer this: is the information correct? Did all information come from the article text?

```{r}
#Loading article text
text = "Maggie Smith, one of the finest British stage and screen actors of her generation, whose award-winning roles ranged from a freethinking Scottish schoolteacher in /“The Prime of Miss Jean Brodie/” to the acid-tongued dowager countess on /“Downton Abbey,/” died on Friday in London. She was 89.

Her death, in a hospital, was announced by her family in a statement issued by a publicist. The statement gave no cause of death.

American moviegoers barely knew Ms. Smith (now Dame Maggie to her countrymen) when she starred in /“The Prime of Miss Jean Brodie/” (1969), about a 1930s girls’-school teacher who dared to have progressive social views — and a love life. Vincent Canby’s review in The New York Times described her performance as /“a staggering amalgam of counterpointed moods, switches in voice levels and obliquely stated emotions, all of which are precisely right./” It brought her the Academy Award for best actress.

She won a second Oscar, for best supporting actress, for /“California Suite/” (1978), based on Neil Simon’s stage comedy. Her character, a British actress attending the Oscars with her bisexual husband (Michael Caine), has a disappointing evening at the ceremony and a bittersweet night in bed.

In real life, prizes had begun coming Ms. Smith’s way in the 1950s, when at 20 she won her first Evening Standard Award. By the turn of the millennium, she had the two Oscars, two Tonys, two Golden Globes, half a dozen Baftas (British Academy of Film and Television Awards) and scores of nominations. Yet she could go almost anywhere unrecognized.

Until /“Downton Abbey./”

That series followed the Earl of Grantham (Hugh Bonneville), his mostly aristocratic family and his troubled household staff at their grand Jacobean mansion as the world around them, between 1912 and 1925, refused to stand still.

After its 2010 British premiere and its 2011 American debut, the show ran six seasons. Its breakout star, from the beginning, was Ms. Smith, playing Lord Grantham’s elderly and still stubbornly Victorian widowed mother, Violet Crawley, the dowager countess. She disapproved of electric lights, was unfamiliar with the word /“weekend/” and never met a person or situation she couldn’t ridicule with withering imperiousness. When her daughter-in-law considered sending a younger relative for a stay in New York, Lady Violet objected: /“Oh, I don’t think things are quite that desperate./”

Suddenly, in her mid-70s, Ms. Smith was a megastar.

/“It’s ridiculous. I’d led a perfectly normal life until ‘Downton Abbey,’ /” she told Mark Lawson at the B.F.I. and Radio Times Festival in 2017, adding later: /“Nobody knew who the hell I was./”

The closest Ms. Smith had come to such visibility was with the Harry Potter movies. She was Minerva McGonagall, the Hogwarts School’s stern but fearless transformation teacher, in seven of the eight films, from /“Harry Potter: The Sorceror’s Stone/” (2001) to /“Harry Potter: The Deathly Hallows Part 2/” (2011).

McGonagall, wearing high-necked Victorian-style gowns, a distinctive Scottish brooch, and upswept hair beneath a tall, black witch’s hat, was a striking onscreen presence. Yet Ms. Smith did not find herself constantly pursued in public, except by children.

/“A lot of very small people kind of used to say hello to me, and that was nice,/” she recalled on /“The Graham Norton Show/” in 2015. One boy carefully asked her, /“Were you really a cat?/”)"
```

```{r}
maggie_smith <- axolotr::ask(
  prompt = paste("Extract the films and TV shows mentioned in the following text into a CSV file. Using a table structure, include the following headers: title,year,role,director,co-stars,awards. Each film or show should be its own row and the headers should be column names. Use N/A for any data you don't know. Just give me the CSV, without a title and no yapping.", text),
  model = "llama-3.1-8b-instant"
)

maggie_smith_csv <- read_csv(maggie_smith)
```

**Answer: The information included is correct, but it appears to not have all the information. For example, Groq only knows to list two Harry Potter movies despite Smith being in more than two movies. There are also examples of shows in the article that didn't make it into the table.**

------------------------------------------------------------------------

## Q2. Helping with code explanation: Your data journalism instructor has given you a block of code but you can't make sense of it. Ask Groq if it can explain what the code does with this UMD course. Is the answer correct?

```{r}
# Loading R code example
r_code_example = 'umd_courses |>
  filter(str_detect(title, "Climate") & seats > 0) |>
  group_by(department) |>
  summarize(classes = n()) |>
  arrange(desc(classes))'
```

```{r}
umd_response <- axolotr::ask(
  prompt = paste("I have a CSV file of all the courses offered at the University of Maryland. I have been given code using the text name r_code_example. Please explain what the code does. Do not dare yap at me. Get to the point.", r_code_example),
  model = "llama-3.1-8b-instant"
)

print(umd_response)

```

**Answer: The answer is correct. This code will show which department offers the most classes with climate in the title and also having available seats. Groq worded it in a bit of a confusing manner, but it is saying the correct thing.**

------------------------------------------------------------------------

## Q3. Helping with code debugging: paste the code block of an answer you had for a previous lab down here and ask Grok if that code is correct, based on the question in the lab. What do you think about its response?

```{r}
#Loading lab question
lab_question = 'Using case_when(), create a column in the Maryland expenses data called spending_location indicating whether each record indicated money spent in Maryland or outside Maryland, based on the address column of the \'maryland_expenses\' dataset. For records that are in Maryland, make the new column\'s value \"In-state\" and for the others, make it \"Out of state\". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe).'


#Paste the code block here, between the quotes. If your code has quotes or single quotes, you have to add a \ before each one so R doesn't break.

your_r_code_lab = "maryland_expenses |>
  mutate(spending_location = case_when(
    str_detect(address, ' Maryland') ~ 'In-state',
    str_detect(address, ' Virginia') ~ 'Out of state',
    str_detect(address, ' New Jersey') ~ 'Out of state',
    str_detect(address, ' New York') ~ 'Out of state',
    str_detect(address, ' Pennsylvania') ~ 'Out of state',
    .default = 'Out of state or no address'
  )) |>
  group_by(spending_location) |>
  summarise(total_amount = sum(amount))


"


```

```{r}
lab_question_three <- axolotr::ask(
  prompt = paste("Is the code I used in the following text correct", your_r_code_lab),
  model = "llama-3.1-8b-instant"
)

print(lab_question_three)

```

**Answer: This was largely helpful because the code we know is correct based on that it is from the answer key. But on top of that, Groq gives some scenarios that could be another way to answer the same question. Of course, we'd have to actually try that to make sure it works, but is always good to have options.**

------------------------------------------------------------------------

## Q4. Brainstorming about strategies for data analysis and visualization: ask Groq to give you ideas about potential news stories from data analyses using [this data set](https://data.montgomerycountymd.gov/Public-Safety/Crash-Reporting-Incidents-Data/bhju-22kf/about_data). You're going to have to describe the dataset so that Groq can make some suggestions. What do you think of its response?

```{r}
#Add code below to talk to Groq and display its response
crash_data <- axolotr::ask(
  prompt = 'Tell me some news stories I can write about a data set of car crashes in Montgomery County, Maryland. The data includes information on where the crash occured, which police agency responded to the crash, what data the crash occured, and how many cars were involved in the crash.',
  model = "llama-3.1-8b-instant"
)

print(crash_data)

```

**Answer: This response provides pretty basic but decent ideas for data analysis. This answer will not give us specific trends in data because Groq has not seen the data itself. It only knows the general information I input. Most of the ideas here are standards including when crashes most often occur, the severity of the crash, where crashes occurred and one-year comparisons. All of those are important stories, but also are stories that the reporter should think of on their own.**

------------------------------------------------------------------------

## Q5. Using AI to write news articles: ask Groq to write a 500-word news article about a car crash at the intersection between Knox Road and Regents Drive. What issues or problems do you see with the output?

```{r}
car_article <- axolotr::ask(
  prompt = "Write me a 500 word news article about a fatal car crash at the intersection between Knox Road and Regents Drive in College Park, Maryland. Please write this story in proper Associated Press news style and a straight news lead. The story should include a quote from the College Park Police Department and the President of the University of Maryland. There should be a quote from Mickey Mouse and former United States President Jimmy Carter.",
  model = "llama-3.1-8b-instant"
)

car_article

```

**Answer: I told Groq to write me a quote from the UMD president. The name it gave for the quotes is not correct the first time I hit run. Then, I kept messing with the prompt and it eventually spit out the correct name. So the issue we need to keep an eye on as journalists is the accuracy of names and facts used.**
